---
title: "Label-free machine learning analysis of patient-derived tumor organoid drug response"
author: "Nolan Ung, Naim Matasci"
date: "07/08/2019"
always_allow_html: yes
output:
  html_document:
    highligh: haddock
    keep_md: yes
    number_sections: yes
    theme: paper
    toc: yes
    toc_depth: 3
    toc_float:
      collapsed: no
      smooth_scroll: no
  html_notebook:
    highligh: haddock
    number_sections: yes
    theme: paper
    toc: yes
    toc_depth: 3
    toc_float:
      collapsed: no
      smooth_scroll: no
  pdf_document: default
---

## MANUALLY RUN THIS FIRST
Run on `r date()`
```{r checkpoint, eval=FALSE, warning=FALSE, include=FALSE}
### *** IMPORTANT *** ###
#This chunck must be run MANUALLY before attempting to knit this document
#This document uses the package checkpoint for package versioning and ensuring reproducibility
#After manually running the checkpoint command the documents can be knit using RStudio

library(checkpoint)
checkpoint("2018-11-27")

```

# Setup
```{r setup, include=FALSE}
dev.off()
knitr::opts_chunk$set(echo = TRUE)
source("functions.R")
input_folder<-file.path(here::here(), "data")
```

# Plotting options
```{r global_plotting_opts, include=FALSE}
Compound_cols = c(
  Media = "hotpink1",
  OXA = "gold2",
  IRI = "mediumpurple3",
  STA = "turquoise3"
  )

LD_cols = c(Live = "gray80", Dead = "gray25")
  
Patient_cols = c(
  `12415` = "yellow3",
  `12527` = "turquoise1",
  `12620` = "tomato",
  `12737` = "purple",
  `12911` = "blue",
  `13154` = "chartreuse4",
  `12620_A` = "tomato",
  `12620_B` = "dodgerblue",
  `12415_A` = "yellow3",
  `12415_B` = "orange"
  )
```

#Load input
```{r load_input, echo=TRUE, include=FALSE}
#Load and clean the input
big_tbl <- clean_input(import_all(file.path(input_folder, "raw")))
big_tbl<- big_tbl %>% mutate(`Row`=paste(`Row`,ExpID,sep="_")) #Ensuring that all the measurements are kept separated
#names(big_tbl) <- make.names(names(big_tbl))

#GT Organoid classification by experts
org_class <- read_csv(file.path(input_folder,"ground_truth", "Organoid_Classification_Responses.csv"))

#GT Vital Dye measurements
gt_i<-import_wb(file.path(input_folder,"ground_truth","20171010_20171017_12620_GT.xlsx"))
gt_i<-gt_i %>% mutate(UID=paste(Row,Column,Field,`Object No in Organoids`,Day,sep='.'))

#GT Imaging measurements
gt_n<-import_wb(file.path(input_folder,"ground_truth","20171010_20171017_12620_GT_DRAQ7.xlsx"))
gt_n<-gt_n %>% mutate(UID=paste(Row,Column,Field,`(DRAQ7) - Object No in Organoids`,Day,sep='.'))
 

#Cell Titer Glo
ctg_1 <- plater::read_plate(
  file = file.path(input_folder,"ground_truth","20170526_12737_IRI_Curve_D3_ES.csv" ),
  well_ids_column = "Wells"     # name to give column of well IDs (optional)
)
ctg_2 <- plater::read_plate(
  file = file.path(input_folder,"ground_truth","20170530_12737_IRI_Curve_CTG_D7_ES.csv" ),
  well_ids_column = "Wells"     # name to give column of well IDs (optional)
)

ctg_3 <-plater::read_plate(
  file = file.path(input_folder,"ground_truth","20170609_12737_IRI_Curve2_D3_ES.csv" ),
  well_ids_column = "Wells"     # name to give column of well IDs (optional)
)

ctg_4<-plater::read_plate(
  file = file.path(input_folder,"ground_truth","20170613_12737_IRI_Curve2_D7_ES.csv" ),
  well_ids_column = "Wells"     # name to give column of well IDs (optional)
)


# Tracked organoids
tracked_organoids <- read.csv(file.path(input_folder, "ground_truth",  "20180122_13154_GT_Tracked-Nulcei_INT.csv"), 
                      stringsAsFactors = FALSE, 
                      encoding = "latin1")

```

# Fig 2B - Heatmap of organoids in media
```{r fig2_b, echo=FALSE, message=FALSE, warning=FALSE, fig.height=11, fig.width=14}
ann_colors <- list(
  LD = LD_cols, 
  Compound = Compound_cols[c("Media","STA")], 
  Patient = Patient_cols[c(  "12415" ,  "12527","12620","12737", "12911","13154" )]
)

mlt_big_tbl_media <-  big_tbl %>%
  melt(
  id.vars = c(
  "Display",
  "Row",
  "Compound",
  "Concentration",
  "Cell Type",
  "LD",
  "Day",
  "ExpID"
  )
  )  %>%
  filter(grepl("Region|Object", variable)) %>%
  filter(Compound == "Media")

map_media_l <- make_map_pdr(fix_vars2(mlt_big_tbl_media))

pheatmap(mat = map_media_l$mat, 
         scale = "row",
         cluster_rows = TRUE,
         border_color = NA,
         annotation_col  = map_media_l$annot, 
         annotation_colors = ann_colors,
         main = '',
         show_colnames = FALSE)
```

#Fig 2C - Organoid growth
```{r fig2c, echo=FALSE, message=FALSE, warning=FALSE, fig.height=5, fig.width=7}

all_tbl <-
  big_tbl %>% filter(ExpID != "20171012_12415_OXA_2016905_0913" ,
  ExpID != "20181003_12620_IRI_DC")
  
  feature <- "Region Area [µm²]"
  subdata <-
  subset_measurements(all_tbl, feature = feature, missing_as_zero = FALSE)
  
  a_curves <-
  subdata %>% mutate(Patient =  `Cell Type`) %>% filter(Compound ==
  "Media") %>% group_by(Patient, Compound, Concentration, Day) %>%    summarise(
  N_wells = sum(!is.na(Live)),
  live_mean = mean(Live, na.rm = TRUE),
  live_sd = sd(Live, na.rm = TRUE)
  )
  
  pd <- position_identity()
  days <- sort(as.numeric(unique(a_curves$Day)))
  
  pl <- ggplot(data = a_curves) +
  aes(x = Day,
  y = live_mean,
  col = Patient) +
  geom_line(lwd = 1) +
  scale_color_manual(values = Patient_cols[c("12415", "12527", "12620", "12737", "12911", "13154")]) +
  
  scale_x_continuous(breaks = days) +
  ylab(label = feature) +
  labs(title = "") +
  theme_linedraw()
  
  plot(pl)
```

#Fig 3A - Expert Ground Truth
```{r Expert_ground_truth, echo=FALSE, message=FALSE,  warning=FALSE, fig.height=5, fig.width=7}
# melt and make variables factors 
org_melt <-
  org_class %>% melt(id.vars = c("Timestamp", "Email Address", "Group"))
  
  org_melt$Group <-
  ordered(
  org_melt$Group,
  levels = c("experimental", "Harmony", "DRAQ7"),
  labels = c("Experts", "Harmony", "DRAQ7")
  )
  
  
  names(org_class) <-
  make.names(names = names(org_class)) %>% str_replace_all(pattern = "^.*\\.\\.|X\\.|\\.$", replacement = "")
  
  majority <-
  lapply(org_class, function(x)
  names(which.max(table(x)))) %>%  as.data.frame() %>%
  mutate(Email.Address = "Majority@stat.com", Group = "Majority")
  
  melt_major <-
  org_class %>% full_join(majority) %>%  melt(id.vars = c("Timestamp", "Email.Address", "Group"))
  melt_major$Group <-
  ordered(
  melt_major$Group,
  levels = c("experimental", "Majority",  "Harmony", "DRAQ7"),
  labels = c("Experts", "Majority",  "ML", "DRAQ7")
  )
  
  melt_major$variable <- factor(
  melt_major$variable,
  levels = c(
  "A.15",
  "A.19",
  "C.41",
  "D.48",
  "F.29",
  "A.22",
  "A.17",
  "B.44",
  "E.49",
  "C.8",
  "D.82",
  "F.53",
  "D.12",
  "B.22",
  "E.82",
  "C.43",
  "F.59",
  "E.12"
  ),
  labels = c(
  "1",
  "2",
  "3",
  "4",
  "5",
  "6",
  "7",
  "8",
  "9",
  "10",
  "11",
  "12",
  "13",
  "14",
  "15",
  "16",
  "17",
  "18"
  )
  )
  
  
  G <-  ggp(melt_major, aes(x = variable, fill = value))
  G + geom_bar(stat = "count",
  position = "stack",
  width = .95)  +
  labs(title = "Expert Ground Truth for Organoid Viability", x = "Organoid ID") + facet_grid(Group ~
  ., scales = "free_y") +
  geom_hline(
  data = data.frame(yint = 4.5, Group = "Experts"),
  aes(yintercept = yint),
  linetype = "solid",
  color = "black"
  ) +
  theme(legend.title = element_blank()) + scale_y_continuous(breaks = c(1, 3, 5, 7, 9)) +
  scale_fill_manual(values = c("#FAA43A", "#5DA5DA"))

  
```

#Fig 3B - Comparison between VitalDye and ML
```{r comp_ML_VD, echo=FALSE, message=FALSE,  warning=FALSE, fig.height=5, fig.width=7}
#Remove cases where multiple objects map to the same organoid in image
en_b<-nrow(gt_n)
gt_n <- gt_n %>% group_by(UID) %>% filter(n()==1)

if((en_b-nrow(gt_n))>0){warning(paste("Removed",removed, "duplicate entries out of",en_b,"total" ))}

n_suffix<-'_VD'
i_suffix='_Imaging'

gt_r<-unify(gt_n,gt_i,n_suffix,i_suffix)
gt_r<-restack(gt_r,'LD',n_suffix,i_suffix)

subdata<-subset_measurements(gt_r,'counts')
nuc_counts<-normalize_data(subdata,metric="proportion")
cm<-matrix(unlist(strsplit(nuc_counts$Compound,'_')),ncol=2,byrow=T)
nuc_counts$Compound<-cm[,1]
nuc_counts$Assay<-cm[,2]
nuc_counts[nuc_counts$Compound=="media","Compound"]<-"Media"
nuc_counts[nuc_counts$Assay=="Imaging","Assay"]<-"ML"

i_aes<-list(cols=c('black','black','red','red'),shapes=c(0,0,1,1),lines=c(0,0,1,1))

plot(plot_c_curves(nuc_counts,'counts',patient="12620",error_bars = F,normalized = T,metric='proportion')+ theme(axis.text = element_text(size = 15)))
```

#Fig 3C - IRI dose response
```{r iri_dr , echo=FALSE, message=FALSE,  warning=FALSE, fig.height=5, fig.width=7}
subdata<-subset_measurements(big_tbl %>% filter(ExpID == '20181012_20170530_12737_IRI'),feature="counts",missing_as_zero = TRUE)
subdata<-make_treatments(subdata)
#subdata %>% group_by(`Cell Type`, Compound, Concentration,Day) %>% filter(`Cell Type`==12415) %>% summarise(n()) %>% View()


norm_counts<-normalize_data(subdata)

n737<-norm_counts %>% filter (`Cell Type` ==12737)

plot(plot_curves(n737,pat="12737",feature="counts",error_bars = T,normalized = T,metric="proportion"))

```

#Fig 3D - CTG Drug response
```{r CellTiter-Glo_plot_12737-1, message=FALSE, echo=FALSE, warning=FALSE,  fig.height=5, fig.width=7}


#calculate average value of ATP = 0 background control and subract from all values
avg_ATP1 <- ctg_1 %>% filter(Compound == 'ATP' & Concentration == 0)  %>% select(Signal) %>% summarise(mean=mean(Signal,na.rm=TRUE))
day_data1 <- ctg_1 %>% mutate(Day = factor(3)) %>% mutate(Signal = Signal - avg_ATP1$mean) %>% filter(!Concentration == .25 | is.na(Concentration))
day_data1$Compound[day_data1$Compound == "media"] <-  "Media"

#calculate average value of ATP = 0 background control and subract from all values
avg_ATP2 <- ctg_2 %>% filter(Compound == 'ATP' & Concentration == 0)  %>% select(Signal) %>% summarise(mean=mean(Signal,na.rm=TRUE))
day_data2 <- ctg_2 %>% mutate(Day = factor(7)) %>% mutate(Signal = Signal - avg_ATP2$mean)
day_data2$Compound[day_data2$Compound == "media"] <-  "Media"

#combine all plates into one DF and filter only relevant Conditions
ctg_data <- rbind(day_data1, day_data2)

ctg_data <- ctg_data %>%  
    filter(Compound != "ATP") %>% 
  group_by(Compound, Concentration, Day) %>% 
  dplyr::summarise(mean = mean(Signal), sd = sd(Signal)) 

#normalize by media values each day
norm_temp <-  ctg_data %>% group_by(Compound, Day) %>% 
  dplyr::summarise(mean = mean(mean)) %>% filter(Compound == "Media")


ctg_data <- ctg_data %>% mutate(norm_mean = ifelse(Day == "3", mean/as.numeric(norm_temp[1,3]), ifelse(Day == "7", mean/as.numeric(norm_temp[2,3]), NA))) %>% mutate(norm_sd = ifelse(Day == "3", sd/as.numeric(norm_temp[1,3]), ifelse(Day == "7", sd/as.numeric(norm_temp[2,3]), NA)))


#arrage colors and order of Conditions
ctg_data$Compound <- factor(ctg_data$Compound, levels = c("STA", "IRI", "Media",  "NOC"))
ctg_data$Concentration <- factor(ctg_data$Concentration)
ctg_data <- make_treatments(ctg_data)

# make aesthetics
cols <- make_aes(ctg_data)$cols
shapes <- make_aes(ctg_data)$shapes
shapek <- c(0:(length(shapes)-1))

#calculate min and max for sd bars
miny <- ctg_data$norm_mean - ctg_data$norm_sd
maxy <- ctg_data$norm_mean + ctg_data$norm_sd

dodge <- position_dodge(.8)

G <- ggp(ctg_data, aes(x = Day, y = norm_mean, fill = Treatment, 
                            group = Treatment,
                            ymin = miny, ymax = maxy, 
                         shape = Compound))
G +  geom_bar(stat = "identity", position  = dodge) + 
  geom_errorbar(width = .3, position = dodge, size = 1) +
  scale_fill_manual(values = cols,  name = "Condition") + 
  scale_x_discrete(breaks=c("3", "7"), 
                   labels=c("3", "7")) +  scale_y_continuous(limits = c(0,2)) +
  labs(title = "", y = "Mean Signal", x = "Day") +
    scale_shape_manual(values = shapes, guide = FALSE) +
    guides(colour = guide_legend(override.aes = list(shape = shapes))) +
  geom_hline(yintercept = 1, linetype = "dashed", alpha = .5) +
  ylim(0,1.5)
  
```

#Fig 4 - Live/dead PDOs based on morphology and texture
```{r heatmap_ld, , message=FALSE, echo=FALSE, warning=FALSE,  fig.height=11, fig.width=14}
#### Combine live and dead, Media and STA #####
mlt_big_L_all <-  big_tbl %>%  filter(LD == "Live") %>% melt(id.vars = c("Display", "Row", "Compound", "Concentration", "Cell Type", "LD", "Day", "ExpID"))  %>% filter(grepl("Region|Object", variable)) %>% filter(Compound %in% c("Media", "STA")) 

mlt_big_L_all <- fix_vars2(mlt_big_L_all)

mlt_big_D_all <-  big_tbl %>%  filter(LD == "Dead") %>%  melt(id.vars = c("Display", "Row", "Compound", "Concentration", "Cell Type", "LD", "Day", "ExpID"))  %>% filter(grepl("Region|Object", variable)) %>% filter(Compound %in% c("Media", "STA"))

mlt_big_D_all <- fix_vars2(mlt_big_D_all)

big_map_all <- bind_rows(mlt_big_D_all, mlt_big_L_all) %>% 
  group_by(ExpID, LD, Compound, Patient, Day, variable) %>% 
  dplyr::summarise(mean_val = mean(value)) %>% 
  spread(variable, mean_val) %>% ungroup() %>% 
  unite(Patient, Day, ExpID, Compound, LD, col = P_D_C_LD_ID, remove = FALSE)

mat_all <- big_map_all %>% column_to_rownames("P_D_C_LD_ID") %>% dplyr::select(-Patient, -Day, -ExpID,  -Compound, -LD) %>%
  as.matrix() %>% t()


annot <- big_map_all %>%  column_to_rownames("P_D_C_LD_ID") %>% dplyr::select(Patient, LD) %>% as.data.frame()

ann_colors_drug <- list(
  LD = LD_cols, 
  Patient = Patient_cols[c(  "12415" ,  "12527","12620","12737", "12911","13154" )]
)

pheatmap(mat = mat_all, scale = "row", 
         cluster_rows = TRUE, 
         border_color = NA,
         annotation_col  = annot,
         annotation_colors = ann_colors_drug,
#         main = "Live and Dead PDOs Media and STA treated"
         main = '',
         show_colnames = FALSE)
         
```


#Fig 5A - Response curves
```{r load_data, message=FALSE, echo=FALSE, warning=FALSE,  fig.height=5, fig.width=7}
feature<-"counts"
subdata<-subset_measurements(big_tbl,feature=feature,missing_as_zero = TRUE)
subdata<-make_treatments(subdata)


subdata<-subdata %>% filter( 
  (Compound == "IRI" & Concentration == 20)   |
  (Compound == "IRI" & Concentration == 40)   |
  (Compound == "OXA" & Concentration == 20)   |
    (Compound == "STA")   |
    (Compound == "Media") 
)

norm_counts<-normalize_data(subdata)

plot(plot_curves(norm_counts %>% filter (`Cell Type` ==12415) %>% ungroup() %>% mutate(`Cell Type`=paste(`Cell Type`,"(Primary)")),pat="12415",feature=feature,error_bars = T,normalized = T,metric="proportion" ))

plot(plot_curves(norm_counts %>% filter (`Cell Type` ==12527)%>% ungroup() %>% mutate(`Cell Type`=paste(`Cell Type`,"(Metastasis)")),pat="12527",feature=feature,error_bars = T,normalized = T,metric="proportion"))

plot(plot_curves(norm_counts %>% filter (`Cell Type` ==12911)%>% ungroup() %>% mutate(`Cell Type`=paste(`Cell Type`,"(Primary)")),pat="12911",feature=feature,error_bars = T,normalized = T,metric="proportion"))

```

#Fig 5B - Dead PDO Features
```{r, message=FALSE, echo=FALSE, warning=FALSE,  fig.height=5, fig.width=14}
drug_ann_colors<-list(Compound=Compound_cols)

drug_tbl <- big_tbl %>% melt(id.vars = c("Display", "Row", "Compound", "Concentration", "Cell Type", "LD", "Day", "ExpID"))  %>% filter(grepl("Region|Object", variable)) %>%
 
  filter((Compound == "OXA" & Concentration == 20) | (Compound == "IRI" & Concentration %in% c(20)) | Compound == "Media" | Compound == "STA" | Compound == "STA 5", Day != 0 ) 

drug_tbl_clean <-fix_vars2(drug_tbl)

drug_l_12415=make_map_pdcc_ld(drug_tbl_clean  %>%  filter(ExpID %in% c( "20181109_20161025_1107_12415", "20171012_12415_OXA_2016905_0913"))  ,"Dead")


pheatmap(mat = drug_l_12415$mat, 
         color = viridis(100), 
         fontsize = 14, 
         labels_col = "",
         scale = "none",
         cluster_rows = FALSE,
         cluster_cols = FALSE,
         border_color = NA,
         annotation_col  = drug_l_12415$annot,
         annotation_colors = drug_ann_colors,
         main = "12415 Dead Only")




drug_l_12527=make_map_pdcc_ld(drug_tbl_clean %>% filter(ExpID %in% c("20171012_12527_20161206_1213")),"Dead")

pheatmap(mat = drug_l_12527$mat, 
         color = viridis(100), 
         fontsize = 14, 
         labels_col = "",
         scale = "none",
         cluster_rows = FALSE,
         cluster_cols = FALSE,
         border_color = NA,
         annotation_col  = drug_l_12527$annot,
         annotation_colors = drug_ann_colors,
         main = "12527 Dead Only")

drug_l_12911=make_map_pdcc_ld(drug_tbl_clean %>% filter(ExpID %in% c( "20171016_12911_CRC_20170314_0321")),"Dead")

pheatmap(mat = drug_l_12911$mat, 
         color = viridis(100), 
         fontsize = 14, 
         labels_col = "",
         scale = "none",
         cluster_rows = FALSE,
         cluster_cols = FALSE,
         border_color = NA,
         annotation_col  = drug_l_12911$annot,
         annotation_colors = drug_ann_colors,
         main = "12911 Dead Only")

```

#Supplementary Fig 2A: Phenotypic and Morphological variance between patients treated with Media

```{r violin_plot, echo=FALSE, message=FALSE, warning=FALSE, fig.height=11, fig.width=14}

mlt_big_tbl <- big_tbl %>%  
  filter(Day == 3) %>% 
  mutate(Row = paste(Row, ExpID, sep = "_")) %>%
  reshape2::melt() %>% 
  filter(grepl("Region|Object", variable)) %>%
  filter(Compound %in% c('Media', 'Control')) %>% 
  mutate(variable = str_replace_all(variable, pattern = "Region.", replacement = ""))

mlt_big_tbl_media <- fix_vars2(mlt_big_tbl)

#cairo_ps(file="SF2A.ps",height=11.0,width=15.0)
G <- ggp(mlt_big_tbl_media, aes(x = Compound, y = value, fill = Patient))
G + geom_violin(trim = FALSE) + 
  facet_wrap(~ variable, scales = "free_y") + 
  labs(title = "Media", y = "Feature Values", x = "") +
  scale_fill_manual(values=Patient_cols[c(  "12415" ,  "12527","12620","12737", "12911","13154" )]) +
  theme(axis.ticks.x = element_blank(), axis.text.x = element_blank())
# dev.off()

```


#Supplementary Fig 2B: Phenotypic and Morphological variance between patients treated with STA
```{r violin_plot_STA, echo=FALSE, message=FALSE, warning=FALSE, fig.height=11, fig.width=19}

## plot features and patients.
mlt_big_tbl <-  big_tbl %>%  
  filter(Day == 3) %>%
  mutate(Row = paste(Row, ExpID, sep = "_")) %>%
  reshape2::melt() %>% filter(grepl("Region|Object", variable)) %>%
  filter(Compound %in% c("STA")) %>%
  mutate(variable = str_replace_all(variable, pattern = "Region.", replacement = ""))

mlt_big_tbl_STA <- fix_vars2(mlt_big_tbl)

#cairo_ps(file="SF2B.ps",height=11.0,width=15.0)

G <- ggp(mlt_big_tbl_STA, aes(x = Compound, y = value, fill = Patient))
  G + geom_violin(trim = FALSE) + 
    facet_wrap(~ variable, scales = "free_y") + 
    labs(title = "Staurosporine", y = "Feature Values", x = "") +
    scale_fill_manual(values=Patient_cols[c(  "12415" ,  "12527","12620","12737", "12911","13154" )]) +
    theme(axis.ticks.x = element_blank(), axis.text.x = element_blank())
#dev.off()

### individual feature scatter plots. ###
#
#filter_tbl <- bind_rows(mlt_big_tbl_media, mlt_big_tbl_STA) %>% filter(variable %in% c("Profile 1/2", "Profile 2/2"))

#G <- ggp(filter_tbl, aes(x = Patient, y = value, fill = LD))
#G + geom_violin() + facet_grid(cols = vars(variable))
```


#Supplementary Fig 6A - Sankey Media 

```{r Sankey_plots,echo=FALSE, message=FALSE, warning=FALSE, fig.height=7, fig.width=9 }

# rename and select important columns and create unique ID
sub_organoids <- tracked_organoids %>% 
  rename(harmony_live = Tracked.Organoids.Selected.Selected...Live, 
         harmony_dead = Tracked.Organoids.Selected.Selected...Dead, 
         DRAQ_live_perObject = Tracked.Organoids.Selected.Selected...Number.of.Nuclei..DRAQ7....per.Object, 
         DRAQ_dead = Tracked.Organoids.Selected.Selected...Dead..DRAQ7.., 
         DRAQ_live = Tracked.Organoids.Selected.Selected...Live..DRAQ7.., 
         DRAQ_INT = Tracked.Organoids.Selected.Selected...Intensity.Nucleus.Mean.Smoothed.Mean) %>% 
  dplyr::select(-starts_with('Tracked')) %>% 
  mutate(unique_ID = paste(Row, Column, Field, Object.ID, sep = '_')) %>%
  mutate(Time..s. = round((((Time..s./60)/60)/24))) %>% dplyr::select(-Timepoint) %>% 
  rename(Timepoint = Time..s.)

# filter for organoids that are tracked across all timepoints
surv_organoids <- sub_organoids %>% group_by(Row, Column, Field, Compound, Object.ID) %>% 
  dplyr::summarise(count = n()) %>% filter(count >=5)

# creat a survival status comun and label each organoid based on its L or dead call in harmony or by DRAQ
surv_organoids_only <- inner_join(sub_organoids, surv_organoids) %>% mutate(surv_status = 0)
surv_organoids_only$surv_status[surv_organoids_only$harmony_live == 1 & surv_organoids_only$DRAQ_live == 1] <- "L-ML_L-VD"
surv_organoids_only$surv_status[surv_organoids_only$harmony_live == 1 & surv_organoids_only$DRAQ_live == 0] <- "L-ML_D-VD"
surv_organoids_only$surv_status[surv_organoids_only$harmony_live == 0 & surv_organoids_only$DRAQ_live == 1] <- "D-ML_L-VD"
surv_organoids_only$surv_status[surv_organoids_only$harmony_live == 0 & surv_organoids_only$DRAQ_live == 0] <- "D-ML_D-VD"


# prepare df to be changed into a matrix
organoids_mat <- surv_organoids_only  %>% mutate(surv_status = as.factor(surv_status)) %>% 
  mutate(surv_status_num = as.integer(surv_status)) %>% filter(Timepoint != 2) %>% 
  dplyr::select(unique_ID, Timepoint, surv_status, surv_status_num, DRAQ_INT,Compound) %>% unique() 



sankey_media<-make_sankey(organoids_mat %>% filter(Compound == "Media"))
sankey_sta<-make_sankey(organoids_mat %>% filter(Compound == "STA"))

my_color <- 'd3.scaleOrdinal() .domain(["type_a", "type_b", "a", "b", "c", "d"]) .range(["#FAA43A","#5DA5DA", "#1E90FF", "#DAA520"])'


# Make the Network
sankeyNetwork(Links = sankey_media$links, Nodes = sankey_media$nodes,  
              Source = "IDsource", Target = "IDtarget",
              Value = "count", NodeID = "name", colourScale = my_color,
              NodeGroup = "group", LinkGroup = "group", fontSize = 20, 
              sinksRight = FALSE,fontFamily = "Helvetica")

sankeyNetwork(Links = sankey_sta$links, Nodes = sankey_sta$nodes,  
              Source = "IDsource", Target = "IDtarget",
              Value = "count", NodeID = "name", colourScale = my_color,
              NodeGroup = "group", LinkGroup = "group", fontSize = 20, 
              sinksRight = FALSE,fontFamily = "Helvetica")



```

#Supplemantary Fig 6B  - Machine Leaning vs. Vital Dye 
```{r MLvVD, echo=FALSE, message=FALSE, warning=FALSE, fig.height=7, fig.width=9 }

sub_data_tbl <- subset_measurements(big_tbl)
sub_data_tbl <- normalize_data(sub_data_tbl, metric = 'proportion')
sub_data_tbl <- make_treatments(sub_data_tbl) 
sub_data_tbl$Day <- factor(sub_data_tbl$Day)
sub_data_tbl$Concentration <- factor(sub_data_tbl$Concentration)
sub_data_tbl$Origin <- "Harmony"
sub_data_tbl$Treatment[sub_data_tbl$Treatment == "STA"] <- "STA_5"

#calculate average value of ATP = 0 background control and subract from all values
avg_ATP1 <- ctg_1 %>% filter(Compound == 'ATP' & Concentration == 0)  %>% select(Signal) %>% summarise(mean=mean(Signal,na.rm=TRUE))
day_data_D <- ctg_1 %>% mutate(Day = factor(3)) %>% mutate(Signal = Signal - avg_ATP1$mean) %>% filter(!Concentration == .25 | is.na(Concentration))

avg_ATP2 <- ctg_2 %>% filter(Compound == 'ATP' & Concentration == 0)  %>% select(Signal) %>% summarise(mean=mean(Signal,na.rm=TRUE))
day_data_E <- ctg_2 %>% mutate(Day = factor(7)) %>% mutate(Signal = Signal - avg_ATP2$mean)

avg_ATP3 <- ctg_3 %>% filter(Compound == 'ATP' & Concentration == 0)  %>% select(Signal) %>% summarise(mean=mean(Signal,na.rm=TRUE))
day_data_F <- ctg_3 %>% mutate(Day = factor(3)) %>% mutate(Signal = Signal - avg_ATP3$mean) 

avg_ATP4 <- ctg_4 %>% filter(Compound == 'ATP' & Concentration == 0)  %>% select(Signal) %>% summarise(mean=mean(Signal,na.rm=TRUE))
day_data_G <- ctg_4 %>% mutate(Day = factor(7)) %>% mutate(Signal = Signal - avg_ATP4$mean)

total_data <- bind_rows(day_data_D, day_data_E, day_data_F, day_data_G)
total_data <- make_treatments(total_data)
total_data$Concentration <- factor(total_data$Concentration)
total_data$Treatment[total_data$Treatment == 'media'] <- 'Media'
total_data$Treatment[total_data$Treatment == 'STA'] <- 'STA_5'
total_data$Compound[total_data$Compound == 'media'] <- 'Media'
total_data$`Cell Type`<-as.factor("12737")
total_data$Origin <- "CellTiterGlo"

data_tbl_join <- inner_join(total_data, sub_data_tbl, by = c("Day", "Treatment", "Cell Type", "Concentration")) 

summ_tbl <- data_tbl_join %>% group_by(Day, Treatment, `Cell Type`) %>% 
  dplyr::summarise(mean_sig = mean(Signal), mean_PA = mean(live_mean) ) %>% 
 group_by(Day) %>% mutate(norm_media = (mean_sig / mean_sig[Treatment == "Media"])) %>% 
  filter(!Treatment %in% c('IRI_40', 'OXA_20', 'OXA_40', "Media", "IRI_0.74", "IRI_2.22", "IRI_6.66", "IRI_0.082","IRI_0.25")) %>% filter(Day %in% c(1,3,7)) 

summ_tbl$Treatment <-  factor(summ_tbl$Treatment, labels = c("IRI 20µM", "IRI 60µM", "STA 5µM"))


G <- ggp(summ_tbl, aes( x = norm_media, y = mean_PA, col = Treatment, shape = Day, group = Treatment))
G  + stat_summary(fun.y = 'mean', size = 4, geom = "point") + 
  geom_smooth(size = 1.5, method = 'glm', se = FALSE) +
  labs(title = "",
       x = "Mean Lumiescence Relative to Media (RLU)",
       y = "Mean Percent Live organoids Relative to Media") +
   scale_color_manual(values = c( "#ED873E", "#DF5C24", "#265DAB"), name="") + scale_x_continuous(limits = c(0,1)) + scale_y_continuous(limits = c(0,1))

```